---
title: "Buoy Project"
author: "Wendy Liang"
date: "2020/9/24"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=FALSE,warning=FALSE,message = FALSE)
```

## 1.Introduction

### overview

We need to find evidence of global warming in the data collected by a single weather buoy in the NOAA National Data Buoy Center. In these dataset, there are 18 standard meteorological variables.ATMP, WTMP and DEWP are temperature variable and YY, MM, DD, hh and mm are time variables. In my opinion, what we need to do is finding certain relationship between time and temperature.

### approach outline

* step1:import and clean the data

* step2:deal with the time variables

* step3:observe the data by visualization

* step4:explore the data by building regression model

* step5:gain the conclusion



## 2.Exploratory Data Analysis

I will display how I organize my work in this part.

### load all the library
```{r,warning=FALSE,results='hide',error=FALSE}
library(lubridate)
library(ggplot2)
library(dplyr)
library(xts)
```

#### import the dataset

The cleaning process of the data is in another R.script.You can find it here [link](https://github.com/wendylzw/Buoy-Projet/blob/master/import%20data.R). 

```{r}
Buoy=read.csv("Buoydata.csv")
```


### Variables

#### Add two Variables

Firstly,I add a column called *season* in the dataset. (spring = 1, summer = 2, autumn = 3, winter = 4)

Secondly, I Use *Lubridate* package to create a time variable in form "year-month-date-hour".

```{r}
Buoy$season=0
Buoy$season[Buoy$MM==1|Buoy$MM==2|Buoy$MM==3]=1
Buoy$season[Buoy$MM==4|Buoy$MM==5|Buoy$MM==6]=2
Buoy$season[Buoy$MM==7|Buoy$MM==8|Buoy$MM==9]=3
Buoy$season[Buoy$MM==10|Buoy$MM==11|Buoy$MM==12]=4
Buoy=filter(Buoy,Buoy$hh==12)
time=make_datetime(year = Buoy$X.YY, month = Buoy$MM, day = Buoy$DD)
Buoy=mutate(Buoy,time=time)

#check season
#filter(Buoy,month(time)==5)
```

#### Describe Variables

Here's the url of the data description [link](https://www.ndbc.noaa.gov/measdes.shtml)

**ATMP** is Air temperature (Celsius).

**WTMP** is Sea surface temperature (Celsius). 

**time** is the date.


### Visulization

#### year average of ATMP

This is a plot showing the average air temperature of each year. In addition, the smooth line also show the positive correlation between year and the temperature. In other word, the temperature dose increase over time.

```{r,warning=FALSE}
#year average of ATMP
plane=group_by(Buoy,X.YY)
avg=summarize(plane,avg_ATMP = mean(ATMP))
ggplot(avg,aes(x=X.YY,y=avg_ATMP))+geom_point()+geom_smooth(method ="lm")+xlab("year")
```

#### season average of ATMP

This plot shows the changing curve of the average air temperature of each season, from 1999 to 2019. In addition, the smooth line also show the positive correlation between year and the temperature.

```{r,warning=FALSE}
#sping
sp=filter(Buoy,Buoy$season==1)
plane1=group_by(sp,X.YY)
avg_sp=summarise(plane1,avg_ATMP = mean(ATMP))
avg_sp=mutate(avg_sp,season=1)
#summer
su=filter(Buoy,Buoy$season==2)
plane2=group_by(su,X.YY)
avg_su=summarise(plane2,avg_ATMP = mean(ATMP))
avg_su=mutate(avg_su,season=2)
#autum
aut=filter(Buoy,Buoy$season==3)
plane3=group_by(aut,X.YY)
avg_aut=summarise(plane3,avg_ATMP = mean(ATMP))
avg_aut=mutate(avg_aut,season=3)
#winter
win=filter(Buoy,Buoy$season==4)
plane4=group_by(su,X.YY)
avg_win=summarise(plane4,avg_ATMP = mean(ATMP))
avg_win=mutate(avg_win,season=4)

r=rbind(avg_sp,avg_su,avg_aut,avg_win)

ggplot(r,aes(x=X.YY,y=avg_ATMP))+geom_line()+geom_smooth(methond="lm")+xlab("year")+ggtitle("average air tem of 4 seasons between 1999 to2019")+ facet_grid(. ~ season)


#if choose special day to represent each season, then...
#spr=filter(Buoy,MM==3,DD==20)
#su=filter(Buoy,MM==6,DD==21)
#aut=filter(Buoy,MM==9,DD==22)
#win=filter(Buoy,MM==12,DD==21)
#Buoy_sea=rbind(spr,su,aut,win)
#plot( xts(spr$ATMP, as.Date(spr$time, format='%Y/%m/%d')),type = 'l',main=' ')

```
In this plot, I find that:

* the temperature in 4 seasons all increases over time

+ the temperature in autumn increases most obviously over time

+ the temperature in spring increases least obviously over   


### Regression Model

#### calculate the time length to now

I don't think time in "ymd" form is suitable for regression model, so I change it to numeric form ---- **to_now**. 

**to_now** means the time length from certain date to now.

**to_now_st** means the standard form of **to_now**

```{r}
now=ymd("2020-09-25")
int=interval(Buoy$time,now)
to_now=time_length(int,unit="day")
Buoy=mutate(Buoy,to_now=time_length(int,unit="day"),to_now_st=(to_now-mean(to_now))/sd(to_now))
```

#### build different model, based on "year"
```{r}
#everyday model
fit1=lm(ATMP~to_now_st,data=Buoy)
summary(fit1)

#year average model
avg=summarise(group_by(Buoy,X.YY),avg_ATMP = mean(ATMP),avg_now = mean(to_now))
fit2=lm(avg_ATMP~avg_now,data=avg)
summary(fit2)

#transform:standard year average model
a=(avg$avg_now-mean(avg$avg_now))/sd(avg$avg_now)
fit3=lm(avg_ATMP~a,data=avg)
summary(fit3)

#transform: log year average model
fit4=lm(avg_ATMP~log(avg_now),data=avg)
summary(fit4)
```

According to the coefficient and $R^2$, I think the second model is the best.

We have $ ATMP=-0.44*time+9.09$ and the $R^2=0.15$.In fact, this model fits no well. But we can know that the ATMP increases over times since the slope is $<0$.

#### build different model, based on "season"

Since "autumn" is the best fitted season through the visualization, I just choose its data build regression model. We can use the same method on other three seasons analysis.

```{r}
#aut
aut=filter(Buoy,Buoy$season==3)
plane3=group_by(aut,X.YY)
avg_aut=summarise(plane3,avg_ATMP = mean(ATMP),avg_now = mean(to_now))
avg_aut=mutate(avg_aut)

#season average model
fit5=lm(avg_ATMP~avg_now,data=avg_aut)
summary(fit5)

#transform: standard season average model
aa=(avg_aut$avg_now-mean(avg_aut$avg_now))/sd(avg_aut$avg_now)
fit6=lm(avg_ATMP~aa,data=avg_aut)
summary(fit6)

#transform: log season average model
fit7=lm(avg_ATMP~log(avg_now),data=avg_aut)
summary(fit7)

```
The $R^2$ of the first two model are both better than the third one, so we can use them to describe the data.(We don't need to use log.)0

We have $ATMP=-0.61*time+18.1$ and the $R^2=0.53$. This "season model" fits much better than "year model". There are stronger positive correlation between time and temperature. We can also know that the autumn ATMP increases over time.

By the way, We can use the same method to analyze other three seasons.I plan to not display in this R.script.



## 3.Conclusions

We can find evidence of global warming from these dataset.







