---
title: "MA678 HW5"
author: "Wendy liang"
date: "10/14/2020"
output:
     pdf_document:
         latex_engine: xelatex

---

```{r setup, include=FALSE,message=FALSE,warning=FALSE,error=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(AER)
library(arm)
library(foreign)
library(ggplot2)
library(tidyverse)
library(rstanarm)
library(VGAM)
```

## 15.1 Poisson and negative binomial regression: 
The folder RiskyBehavior contains data from a  randomized trial targeting couples at high risk of HIV infection. The intervention provided  counseling sessions regarding practices that could reduce their likelihood of contracting HIV.  Couples were randomized either to a control group, a group in which just the woman participated,  or a group in which both members of the couple participated. One of the outcomes examined  after three months was “number of unprotected sex acts.”  

### a) 
Model this outcome as a function of treatment assignment using a Poisson regression. Does  the model fit well? Is there evidence of overdispersion?  

Firstly, I make a sumary.

To summarize:

- `sex` is the sex of the person, recorded as "man" or "woman" here
- `couples` is an indicator for if the couple was counseled together
- `women_alone` is an indicator for if the woman went to counseling by herself
- `bs_hiv` indicates if the individual is HIV positive
- `bupacts` is the number of unprotected sex acts reported as a baseline (before treamtnet)
- `fupacts` is the number of unprotected sex acts reported at the end of the study

```{r}
#summary
risk <- read.csv("risky.csv",header=T)
risk$fupacts = round(risk$fupacts)
risk$couples=factor(risk$couples)
risk$women_alone=factor(risk$women_alone)
summary(risk)

```

Then, I use poisson model to fit.
```{R}
#possion regresison
fit1=glm(fupacts~women_alone,family = poisson,data=risk)
summary(fit1)

#check over
portion_factor = 13064/432
portion_factor

#plot(risk$women_alone,risk$fupacts)
#curve(exp(coef(fit1)[1]+coef(fit1)[2]*x),add=TRUE)
```

Although the `woman_alone` is statistically significant, the [poisson model is poor. Since the portion factor = Residual/degrees >> 1, so this model may has overdisperision.

### b) 
Next extend the model to include pre-treatment measures of the outcome and the additional  pre-treatment variables included in the dataset. Does the model fit well? Is there evidence of  overdispersion?  

```{r}
fit2=glm(fupacts~women_alone+sex+bupacts+couples+bs_hiv,family=poisson,data=risk)
summary(fit2)

#residual plot
plot(predict(fit2),residuals(fit2),xlab="Predicted",ylab="Residuals")
abline(a=0,b=0)

#overdispersion test
#[link]https://www.sciencedirect.com/science/article/abs/pii/030440769090014K
dispersiontest(fit2)
```
- The residual increases as the predicted values increase.

- This model has improved but it's still not good.

- I use function `dispersiontest()` and find this model has high overdispersion of 29.65.


### c) 
Fit a negative binomial (overdispersed Poisson) model. What do you conclude regarding  effectiveness of the intervention?

```{r}
fit3=glm(fupacts~women_alone+sex+bupacts+couples+bs_hiv,family=quasipoisson,data=risk)
display(fit3)
```
The intervention decreased the number of unprotected sex acts `fupacts`. When only women participate in, the decrease is  $1-e^{-0.66} = 1-0.51685=48.31% $. When both men and women participate in, the decrease is  $1-e^{-0.41} = 1-0.66365=33.63% $

### d) 
These data include responses from both men and women from the participating couples.  Does this give you any concern with regard to our modeling assumptions? 

In fact, the predictor `women_alone` and `couples` are not independent.
  
  
  
  
## 15.3 Binomial regression: 
Redo the basketball shooting example on page 270, making some changes:  

### (a) 
Instead of having each player shoot 20 times, let the number of shots per player vary, drawn  from the uniform distribution between 10 and 30.  

```{r}
N <- 100
height <- rnorm(N, 72, 3)
p <- 0.4 + 0.1*(height - 72)/3
n <- round(runif(N,10,30)) #I forget round()
y <- rbinom(N, n, p)
data <- data.frame(n=n, y=y, height=height)

fit3_a <- stan_glm(cbind(y, n-y) ~ height, family=binomial(link="logit"),data=data,refresh=0)
print(fit3_a)
```

### (b) 
Instead of having the true probability of success be linear, have the true probability be a  logistic function, set so that Pr(success) = 0.3 for a player who is 5'9" and 0.4 for a 6' tall  player. 

```{r}
N <- 100
height <- rnorm(N, 72, 3)
p<-ifelse(height>72,0.4,0.3)
n <- round(runif(N,10,30))
y <- rbinom(N, n, p)
data <- data.frame(n=n, y=y, height=height)
fit3_b <- stan_glm(cbind(y, n-y) ~ height, family=binomial(link="logit"),data=data,refresh=0)
print(fit3_b)
```
  
  
  
  
## 15.7 Tobit model for mixed discrete/continuous data: 
Experimental data from the National Supported  Work example are in the folder Lalonde. Use the treatment indicator and pre-treatment variables  to predict post-treatment (1978) earnings using a Tobit model. Interpret the model coefficients. 

This dataset description is [link]http://www.stat.columbia.edu/~gelman/arm/examples/lalonde/NSW.vars.final.doc.

I gain help from [link]https://github.com/IamGianluca/arm/blob/master/ch6/arm_ch6p5.ipynb

```{r}
nsw = read.dta("NSW_dw_obs.dta")

# variables as factor
nsw$sample = factor(nsw$sample, labels=c("NSW", "CPS", "PSID"))
nsw$black = factor(nsw$black)
nsw$hisp = factor(nsw$hisp)
nsw$nodegree = factor(nsw$nodegree)
nsw$married = factor(nsw$married)
nsw$treat = factor(nsw$treat)
nsw$educ_cat4 = factor(nsw$educ_cat4, labels=c("less than high school", "high school", "sm college", "college"))
```


When I observe the data file, I find out many same values. So I make a histogram to see data distribution.
```{r,echo=FALSE}
#plot re78
ggplot(data=nsw, aes(x=re78)) + geom_histogram(binwidth=(range(nsw$re78)[2] - range(nsw$re78)[1])/150)
```
From this plot, I find two peak count concentrated on value 0 and 25564.67. This data `re78` may be censored data for some reasons. So there are two subsets: 1) 0 to 25564.67; 2) 25564.67 to the max value.

```{r}
#split indicator
nsw$ind=ifelse(nsw$re78>=25564.669921875,1,0)
nsw$ind=factor(nsw$ind,labels=c("up","low"))

#regression model to predict ind
fit4=glm(ind~age+educ+re75+black+married,family=binomial(link="logit"),data=nsw)
#summary(fit4)
pre.ind=predict(fit4,nsw,type = "response")
real.ind=ifelse(nsw$re78>=25564.669921875, 1, 0)
error.rate=mean((pre.ind>0.5 & real.ind==0)|(pre.ind<0.5 & real.ind==1))
print(error.rate)

```

The error rate of the prediction of indicator is 15.88%, which means this ind regression model `fit4` fit well.So I will use predicted indicator `pre.ind` to replace the `real.ind`.

```{r,warning=FALSE,message=FALSE}
library(VGAM)
#create two subsets
nsw$pre.ind=pre.ind
sub1=nsw[nsw$pre.ind<0.5, ]
sub2=nsw[nsw$pre.ind>=0.5, ]

# two tobit regression
fit5_1=vglm(re78 ~ age + educ + re75,tobit(Lower=0, Upper=25563),data=sub1)
#summary(fit5_1)
pre_1=predict(fit5_1,nsw)

fit5_2=vglm(re78 ~ age + educ + re75,tobit(Lower=25564, Upper=max(nsw$re78)),data=sub2)
#summary(fit5_2)
pre_2=predict(fit5_2,nsw)

pre.re78=ifelse(pre.ind<0.5,
                ifelse(pre_1>0,pre_1,0),
                pre_2)
real.re78=nsw$re78
y=data.frame(cbind(real.re78,pre.re78))
ggplot(data=y)+geom_histogram(aes(x=real.re78),fill="red",binwidth = (range(real.re78)[2]-range(real.re78)[1])/150)+geom_histogram(aes(x=pre.re78),fill="blue",binwidth = (range(pre.re78)[2]-range(pre.re78)[1])/150)
```
  
  
  
  
## 15.15 Summarizing inferences and predictions using simulation: 
Exercise 15.7 used a Tobit model to  fit a regression with an outcome that had mixed discrete and continuous data. In this exercise  you will revisit these data and build a two-step model: 
(1) logistic regression for zero earnings  versus positive earnings, and 
(2) linear regression for level of earnings given earnings are positive. 

Compare predictions that result from each of these models with each other. 

```{r}
#logistic
nsw$earningind1=ifelse(nsw$re78==0,0,1)
fit9=glm(earningind1~age+educ+re75+married+black,data=nsw,family = binomial(link = "logit"))
display(fit9)

#linear regression
nsw=filter(nsw,re78!=0)
nsw$level=ifelse(nsw$re78>=25564,1,0)
fit10=glm(level~age+educ+re75+married+black,data=nsw)
display(fit10)

```
  
  
  
  
## 15.8 Robust linear regression using the t model: 
The folder Congress has the votes for the Democratic and Republican candidates in each U.S. congressional district in 1988, along with the parties’vote proportions in 1986 and an indicator for whether the incumbent was running for reelection  in 1988. For your analysis, just use the elections that were contested by both parties in both  years.  

Inc means the incumbency status in district i in election t, coded as 1 for Democratic incumbents, 0 for open seats, −1 for Republican incumbents)

```{r}
congress = read.csv("congress.csv")
congress88 <- data.frame(vote=congress$v88_adj,pastvote=congress$v86_adj,inc=congress$inc88)
```

### (a) 
Fit a linear regression using stan_glm with the usual normal-distribution model for the  errors predicting 1988 Democratic vote share from the other variables and assess model fit.  

```{r}
congress88$inc=factor(congress88$inc)
fit6=stan_glm(vote~pastvote+inc,data=congress88,refresh=0)
print(fit6)

#predict
congress88$vote_hat=coef(fit6)[1]+coef(fit6)[2]*congress88$pastvote+coef(fit6)[3]*ifelse(congress88$inc==0,1,0)+coef(fit6)[4]*ifelse(congress88$inc==1,1,0)+rnorm(length(congress88$vote),0,1)

#predict vs real plot
ggplot(data=congress88)+geom_histogram(aes(x=vote,fill="vote"))+geom_histogram(aes(x=vote_hat,fill="vote_hat"))

#error 
error_vote=mean(congress88$vote-congress88$vote_hat)
print(error_vote)

```
This model fit well since the mad_sd is low and the error of 0.08485 is small.


### (b) 
Fit the same sort of model using the brms package with a t distribution, using the brm function with the student family. Again assess model fit.  

```{r, warning=FALSE,error=FALSE,message=FALSE}
library(brms)

fit7=brm(vote~pastvote+inc,data=congress88,family = student,refresh=0)
summary(fit7)

#predict
congress88$vote_hat_brm=predict(fit7)[1]

#error 
error_vote=mean(congress88$vote_hat_brm-congress88$vote)
print(error_vote)
```

### (c) 
Which model do you prefer? 

Comparing the error rate, there are not too much difference between these two model.
  
  
  
  
## 15.9 Robust regression for binary data using the robit model: 
Use the same data as the previous  example with the goal instead of predicting for each district whether it was won by the  Democratic or Republican candidate.  

### (a) 
Fit a standard logistic or probit regression and assess model fit.  

```{r}
congress88$win=ifelse(congress88$vote>0.5,1,0)
fit159_a=glm(win~pastvote+inc,data = congress88,family = binomial(link = "probit"))
display(fit159_a)
```

### (b) 
Fit a robit regression and assess model fit.  

```{r,eval=FALSE,results=FALSE}
fit159_b=glm(win~pastvote+inc,data = congress88,family = binomial(link = gosset(2)))
display(fit159_b)
```

### (c) 
Which model do you prefer? 

Comparing their null deviance and residual deviance, there are not too much difference between these two model. are close to each other.
  
  

  
  
## 15.14 Model checking for count data: 
The folder RiskyBehavior contains data from a study of  behavior of couples at risk for HIV; see Exercise 15.1. 

### (a) 
Fit a Poisson regression predicting number of unprotected sex acts from baseline HIV status. Perform predictive simulation to generate 1000 datasets and record the percentage of  observations that are equal to 0 and the percentage that are greater than 10 (the third quartile  in the observed data) for each. Compare these to the observed value in the original data.  

```{r}
fita <- glm(fupacts ~ women_alone+sex+couples+bs_hiv, family=poisson, data=risk)

set.seed(12)
women_alone=factor(sample(c(0,1),1000,replace = T))
bs_hiv=sample(c("positive","negative"),1000,replace = T)
sex=sample(c("woman","man"),1000,replace = T)
couples=factor(sample(c(1,0),1000,replace = T))
newdata=data.frame(women_alone,sex,couples,bs_hiv)

y=predict(fita,newdata)
equal0=(sum(exp(y)==0)/1000)
great10=(sum(exp(y)>10)/1000)

equal0
great10
```

### (b) 
Repeat (a) using a negative binomial (overdispersed Poisson) regression.  

```{r}
fitb=glm.nb(fupacts~women_alone+sex+couples+bs_hiv,data=risk)

set.seed(12)
women_alone=factor(sample(c(0,1),1000,replace = T))
bs_hiv=sample(c("positive","negative"),1000,replace = T)
sex=sample(c("woman","man"),1000,replace = T)
couples=factor(sample(c(1,0),1000,replace = T))
newdata=data.frame(women_alone,sex,couples,bs_hiv)

y=predict(fitb,newdata)
equal0=(sum(exp(y)==0)/1000)
great10=(sum(exp(y)>10)/1000)

equal0
great10


```

###(c) 
Repeat (b), also including ethnicity and baseline number of unprotected sex acts as inputs. 

```{r}
fitc = glm.nb(fupacts ~ women_alone+sex+couples+bs_hiv+bupacts, data=risk)
newdata$bupacts=sample(c(0,max(risk$bupacts)),1000,replace=T)
y=-predict(fitc,newdata)

y=predict(fitc,newdata)
equal0=(sum(exp(y)==0)/1000)
great10=(sum(exp(y)>10)/1000)

equal0
great10

```






