---
title: "615 Final Project"
author: "Wendy Liang"
date: "12/13/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,fig.align = 'center',fig.height=3,fig.width=6)
library(lubridate)
library(dplyr)
library(stringr)
library(XML)
library(RCurl)
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(kableExtra)
library(arm)
library(readr)
library(ggthemr)
```

# Data Clean
## Load Data

```{r}
#details=read.csv("tmdb_data.csv")
#toprate=read.csv("top_rate.csv")
```

## Join Data
```{r}
#raw=left_join(toprate,details,"id")
#write.csv(raw,"raw.csv")
#write.csv(main,"top_rate_full.csv")
main=read.csv("top_rate_full.csv")
```


##  Extract Values with Regular Expressions 
Variables `genres`,`production_companies`,`production_contries`,`spoken_languages` are in JSON format. Transformation of these variables is done in Excel. In this part, I want to show how to clean this kind of raw data in R. So, I just take one variables for example.

I use `str_extract`, `str_c`, `str_split` functions in  `stringr` package to transform these `spoken languages` in R. 
```{r}
#observe
language_num <- str_count(main$spoken_language, "\\)")
lan_max = max(unique(language_num))
#max=10

#split into 10 cols
language_df = str_split_fixed(main$spoken_language, "\\,list", lan_max)
language_df = as.data.frame(language_df,stringsAsFactors = FALSE)

#regular expression
language_name <- as.data.frame(sapply(language_df, function(x) str_remove_all(x, "list")), stringsAsFactors = F)

language_name <- as.data.frame(sapply(language_name, function(x) str_remove_all(x, "\\(english_name = ")), stringsAsFactors = F)

language_name <- as.data.frame(sapply(language_name, function(x) str_remove_all(x, '\\)')), stringsAsFactors = F)
language_name <- as.data.frame(sapply(language_name, function(x) str_remove_all(x, '\\"')), stringsAsFactors = F)


#combine Vi into one col
language_name[is.na(language_name)] <-""

main$spoken_language=str_c(language_name$V1,language_name$V2,language_name$V3,language_name$V4,language_name$V5,language_name$V6,language_name$V7,language_name$V8,language_name$V9,language_name$V10,sep=",")
```

```{r}
#observe
company_num <- str_count(main$production_companies, "list")
com_max = max(na.exclude(unique(company_num)))
#max=26

#split into 8 cols
company_df = str_split_fixed(main$production_companies, "\\,list", com_max)
company_df = as.data.frame(company_df,stringsAsFactors = FALSE)

#regular expression
company_df <- as.data.frame(sapply(company_df, function(x) str_remove_all(x, "list")), stringsAsFactors = F)
company_df <- as.data.frame(sapply(company_df, function(x) str_remove_all(x, '"')), stringsAsFactors = F)
company_df <- as.data.frame(sapply(company_df, function(x) str_remove_all(x, "(?<=\\,\\s).+(?=\\})")), stringsAsFactors = F)
company_df <- as.data.frame(sapply(company_df, function(x) str_remove_all(x, "\\{name\\s\\=\\s")), stringsAsFactors = F)
company_df <- as.data.frame(sapply(company_df, function(x) str_remove_all(x, "\\,\\s\\}")), stringsAsFactors = F)
#for one string
company_df <- as.data.frame(sapply(company_df, function(x) str_remove_all(x, "(?<=\\,).+")), stringsAsFactors = F)
company_df <- as.data.frame(sapply(company_df, function(x) str_remove_all(x, "\\,")), stringsAsFactors = F)

company_df <- as.data.frame(sapply(company_df, function(x) str_remove_all(x, "\\(")), stringsAsFactors = F)
company_df <- as.data.frame(sapply(company_df, function(x) str_remove_all(x, "\\}")), stringsAsFactors = F)
company_name=company_df

#combine Vi into one col
company_name[is.na(company_name)] <- ""

main$production_companies=str_c(company_name$V1,company_name$V2,company_name$V3,company_name$V4,company_name$V5,company_name$V6,company_name$V7,company_name$V8,company_name$V9,company_name$V10,company_name$V11,company_name$V12,company_name$V13,company_name$V14,company_name$V15,company_name$V16,company_name$V17,company_name$V18,company_name$V19,company_name$V20,company_name$V21,company_name$V22,company_name$V23,company_name$V24,company_name$V25,company_name$V26,sep=",")
```


## Add two variables
```{r}
main$release_year=year(main$release_date)
main$release_month=month(main$release_date)

#write.csv(main,"movie_newest.csv")
```
