---
title: "Berry Data Analysis Report"
author: "Wendy Liang"
output:
  pdf_document: 
     latex_engine: xelatex
subtitle: Fall 2020 MA615 Assignment 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE,fig.align="center")
library(ggplot2)
library(dplyr )
library(magrittr)
```


## Data Description
This report is a limited data exploratory analysis of Berry data from the USDA database selector: <https://quickstats.nass.usda.gov>. The data were <https://quickstats.nass.usda.gov/results/D416E96E-3D5C-324C-9334-1D38DF88FFF1> stored online and then downloaded as a CSV file.

The Berry data displays the survey statistics of three kinds berries based on the simple filtering rules of What, Where, and When.<https://quickstats.nass.usda.gov/param_define> and <https://quickstats.nass.usda.gov/src/glossary.pdf> describe different variables of Berry. And My task is to explore these variables and my goal is to support the agricultural research team for analysis and modeling.

  
  
## Method Overview
There are three steps of my work:

- step1: Data Cleaning and Organizing

- step2: Exploratory Data Analysis

- step3: Shiny App

Analysis was conducted in R Studio (Version 1.3.1073). Data cleaning was performed using package `stringr`  and package `magrittr`. EDA was conducted using package `ggplot2` and `dplyr`, creating shiny app were conducted using package `shiny`. In addition, this report was compiled by package `Knitr`.

  
   
## step1: Data Cleaning
In this step, the most important thing is to separate certain chr variables.

Firstly, I use `separate` function on `Data Item` and delete the `Berries Name`.

```{r eval=FALSE}
#### seperate Measure and Berry
ag_data %<>% separate(`Data Item`,c("Berry","Other"),sep = "BERRIES")

#### delete Berry Category
ag_data %<>% select(-Berry)
head(ag_data)
```

Secondly, I use `separate` function with parameter `sep=""` to separate measurement variable, such as "$ / LB".

```{r eval=FALSE}
#### Create Measure Method
ag_data %<>% separate(`Other`,c("lab1","Measure"),sep="MEASURED IN")
#check
head(ag_data)
ag_data%>%summarize_all(n_distinct)
unique(ag_data$Measure)
```

Then, I use functions in `stringr` package to deal with the data. Instead of splitting characters into different column in Prof's work, I create new column. I use `str_extract` function to extract my goal patterns from each string and save them as new variables (new column). In this way, I create `Type`,`Production` and `Marketing` variables.

```{r eval=FALSE}
#### Create Type, Production, Marketing, Domain,Chemi_family, Materials
ag_data$Type=str_extract(ag_data$lab1,pattern = "(BEARING)|(TAME)|(WILD)")
ag_data$Production=str_extract(ag_data$lab1,pattern = "PRODUCTION")
ag_data$Marketing=str_extract(ag_data$lab1,pattern ="(ACRES HARVESTED)|
(YIELD)|(FRESH MARKET)|
(NOT SOLD)|(PROCESSING)|(UTILIZED)|
(APPLICATIONS)|(TREATED)|(NOT HARVESTED)") 
```
  
Last but not least, I saperate` Domain` and `Domain Category` as the same as dealing with `Measurement`.  
  
```{r eval=FALSE}
ag_data %<>% separate(`Domain`,c("Domain","Chemi_family"),sep=",")
ag_data %<>% separate(`Domain Category`,c("lab2","Materials"),sep="[(]")
ag_data %<>% separate(`Materials`,c("Materials","lab3"),sep="[)]")
#check
head(ag_data)
unique(ag_data$Materials)
```

Now, I finish my data cleaning. The following is the result data frame which I export as a csv file. 
  
```{r echo=FALSE}
head(read.csv("new_berry.csv"))
```

  
  
## Step2: Explorary Data Analysis

### Explore Category Variable

This is a histogram plot for `Commodity`, divided by `State`.

```{r echo=FALSE}
## histogram for Commodity, divided by State
berry=read.csv("new_berry.csv")
berry=data.frame(berry)
# organize the Value as numeric, and delete the NA and D 
berry=berry %>% filter(Value != "(NA)") %>% filter(Value != "(D)") 
berry$Value=as.numeric(berry$Value)

ggplot(berry)+geom_histogram(aes(x=Commodity,fill=State),stat = "count")+theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
  
This is a histogram plot for `Chemi_family`, divided by `Commodity`.

```{r echo=FALSE}
## histogram for Chemi, divided by Commidty
ggplot(berry)+geom_histogram(aes(x=Chemi_family,fill=Commodity),stat = "count")+theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
  
This is a pie plot for `Type`, so we can observe 88.06% Berries are BEARING type.
  
```{r echo=FALSE}
## pie for Type, and we can use the same way to plot other pie
T=table(berry$Type)
percentT=prop.table(T)
tbT=cbind(T,percentT)
piepercentT = paste(round(100*tbT[,2],digits = 2), "%")
nameT=c("BEARING","TAME","WILD")
pie(T ,labels=piepercentT ,col=c("purple","green","white"))
legend("topright", nameT,cex=0.8,fill=c("purple","green","white"))

```

  
  
### Explore Contious Variable `Value` 

I'm gonna explore the independent variable `Value` of Strawberry dataset.

1. **Overview of `Measurement` and `Value`**

    + There are 12 types of measurements for strawberry in the dataset.

    + This is a plot for count of each measurement.There are 272 *PCT of area bearing* and 7 *$ / Ton*, which are the max and min along all measurements.  
    
    + This are several boxplots for `Valu`e, divided by `Measurement`.According to the plot, I find the values of different measurements are vary widely. 

```{r echo=FALSE}
### I choose *Strawberries*
sberry=berry%>% filter(Commodity=="STRAWBERRIES")

####################### Overview of Measurement and Value #################
### count of each measurement
mea_sum = sberry %>%
  group_by(Measure)%>%
  summarize(
    Count=n(),
    Mean=round(mean(Value),2)
  )

### plot for count of each measurement
ggplot(mea_sum,aes(x=Measure,y=Count))+
  geom_bar(stat='identity')+
  ggtitle("Measurements of Strawberry")+
  geom_text(aes(label=Count,y=Count+10),size=3) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r echo=FALSE}
### boxplot for Value, divided by Measure
ggplot(sberry, aes(x = Measure, y= Value)) +
  geom_boxplot()+
  xlab("Measurement") + 
  ylab("Mean Value of Each Measurement")+
  facet_wrap(~Measure,scales="free")
```

2. `Value` of three different `Measurement`

    + I choose three different measurements to describe three characters of strawberry. `NUMBER` refers to number, `$ /CWT` refers to price, and `TONS` refers to weight. I calculate the sum of `Value` in each year when exploring number and weight. And I calculate the mean of `Value` in each year when exploring price. 

    + Using `ggplot2` package,I make three point and line plots. According to these plots, we can find the growing trend and value difference.  

```{r echo=FALSE}
######## 1.Measurement == NUMBER ---- Production Number #########
### count of each measurement
sberry_number = sberry %>% 
  filter(Measure=="NUMBER") %>% 
  group_by(State,Year) %>%
  summarise(Marketing=Marketing,Sum=sum(Value))

## plot for Sum of Value in NUMBER measurement
ggplot(sberry_number)+
  geom_line(aes(x=Year,y=Sum,col=State),alpha=0.3)+
  geom_point(aes(x=Year,y=Sum,col=State))+
  xlab("Year") + 
  ylab("Sum of Value") + 
  ggtitle("Number: Value in NUMBER measurement")
```

    
```{r echo=FALSE}
############### 2.Measurement == $ / CWT ---- Price ################
### count of each measurement
sberry_cwt_price = sberry %>% 
  filter(Measure=="$ / CWT") %>% 
  group_by(State,Year) %>%
  summarise(Marketing=Marketing,Mean_price=mean(Value))

## plot for Sum of Value in $ /CWT measurement
ggplot(sberry_cwt_price)+
  geom_line(aes(x=Year,y=Mean_price,col=State),alpha=0.3)+
  geom_point(aes(x=Year,y=Mean_price,col=State,size=Mean_price), alpha=0.5)+
  xlab("Year") + 
  ylab("Mean of Value") + 
  ggtitle("Price: Value in $ / CWT")
```

  
```{r echo=FALSE}
############### 3.Measurement == CWT ---- Weight ################
### count of each measurement
sberry_cwt = sberry %>% 
  filter(Measure=="CWT") %>% 
  group_by(State,Year) %>%
  summarise(Marketing=Marketing,Sum=sum(Value))

## plot for Sum of Value in CWT measurement
ggplot(sberry_cwt)+
  geom_point(aes(x=Year,y=Sum,col=State),size=3)+
  geom_line(aes(x=Year,y=Sum,col=State),alpha=0.3)+
  geom_point(aes(x=Year,y=Sum,col=State),size=3)+
  xlab("Year") + 
  ylab("Sum of Value") + 
  ggtitle("Weight: Value in CWT")
```

  
* There are some observations:

    + number of strawberry increases from 2016 to 2019 year

    + California has the most number of strawberry

    + Price varies widely from states and years

    + The strawberry in New York is the most expensive while California is cheapest.

    + The weight of strawberry in California and Michigan decreased a lot in 2019



#### `Value` of different `Marketing`

I just display one situation in the report ---- `Marketing`== "APPLICATION" meanwhile `Measurement` == "LB". This kind of plot indicate the `Value` distribution of different chemical families.

\includegraphics{7.chemmea.png}



  
## Step3: Shiny App
Here's the APP link: <https://wendy-liang.shinyapps.io/Berry_EDA/?_ga=2.148296823.1134128627.1603185041-870852303.1603185041>



  
## Conclusion

Based on the Exploratory Data Analysis, I come up with some personal conclusions. Firstly, California is a suitable place for strawberry to grow and New York is a profitable place for growers. Secondly, there are correlations between chemical family and value in each measurement. Thirdly, strawberries are grown more with time while prices fluctuate a lot each year.

This analysis certainly contains its limitations. There are a large number of NA in the dataset. My analysis only accounts one kind of berries. Further exploration and modeling could include all these berries -- blueberry, raspberry and strawberry.


  
  
## Reference
Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source
  Software, 4(43), 1686, https://doi.org/10.21105/joss.01686
  
H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New
  York, 2016.

Yihui Xie (2020). knitr: A General-Purpose Package for Dynamic Report
  Generation in R. R package version 1.29.


  


